---
title: 'iCod: Confrontation with sam'
author: "Einar Hjörleifsson"
date: "10/07/2014"
output:
  html_document:
    fig_height: 7
    fig_width: 9
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
---

# Input

## File structure

The input data are in Lowestoft format stored in the *data*-directory:
```{r}
dir("data")
dir.data     <-  file.path("data")
```


## FLR objects

Load the FLCore package
```{r, message=FALSE, warning=FALSE}
library(FLCore)
```
Read the input data into an *FLStock* object:
```{r}
stk <- readFLStock(file.path(dir.data, "index.txt"),no.discards=TRUE)
units(stk)[1:17] <- as.list(c(rep(c("tonnes","thousands","kg"),4), "m","NA","f","NA","NA"))
landings(stk)    <- landings(stk) * 1000
landings.wt(stk) <- landings.wt(stk)/1000
stock.wt(stk)    <- stock.wt(stk) / 1000
stk@catch.n      <- stk@landings.n
stk@catch        <- stk@landings
stk@catch.wt     <- stk@landings.wt
#Set fbar
range(stk)[c("minfbar","maxfbar")] <- c(5,10)
#Set plus group
range(stk)[3] <- 0
iCod.stk <- stk
```
Read the tuning data into an *FLIndices* object:
```{r}
tun   <- readFLIndices(file.path(dir.data, "fleet.txt"))
tun   <- lapply(tun,function(idx) {
                 idx@type        <- 	"number"
          		    idx@index.var[]  <-	1
                  #idx@range["plusgroup"] <- 10  ### NEED TO DOUBLE CHECK
          		return(idx)})
```

# SAM assessment
Note the *FLSAM* package is a forked one from *flr* to *einarhjorleifsson* on github. Hence to install the modified verssion one does:
```{r, eval=FALSE}
require(devtools)
install_github("FLSAM","einarhjorleifsson")
```

```{r, message=FALSE, warning=FALSE}
library(FLSAM)
```

Firstly lets just use the time series since 1985 (start of the SMB survey):
```{r}
stk <- window(iCod.stk,start = 1985)
```

## Setup
Setup an *FLSAM.control* object based on the FLStock and FLIndices objects suppplied:
```{r}
ctr <- FLSAM.control(stk,tun)
```
The above only sets up the structure, but all parameters are still "unset". Lets fill the stuff with something just to get us going:
```{r}
ctr@plus.group <- FALSE
# Survey catchability
ctr@catchabilities["SMB",]  <- c(1,2,3,4,5,6,7,8,8,8,NA,NA,NA,NA)
# Survey power (Should only be up to age 5)
ctr@power.law.exps["SMB",] <-  c(1,2,3,4,5,NA,NA,NA,NA,NA,NA,NA,NA,NA)

# No fishing age 1 and 2:
ctr@f.vars["catch",1:2] <- NA
# Fishing mortality states: for ages 3:10 are free
#  for 11:14 assume the same as age 10
ctr@states["catch",] <- c(NA,NA,1:8,8,8,8,8)

#Bind the observation variances
ctr@obs.vars["catch",] <- c(NA,NA,1,2,2,2,2,2,2,3,3,3,3,3)
ctr@obs.vars["SMB",]   <- c(rep(4,10),NA,NA,NA,NA)

#Correlated Random walks for fishing mortalities - Default = FALSE = independent)
ctr@cor.F <- FALSE

#Set the variances. Separate variance for recruitment and plus group
ctr@logN.vars[]      <- c(1,rep(2,dims(stk)$age-1))
#ctr_trial01 <- ctr # a backup
```

## Assessment
```{r, eval=FALSE}
sam   <-  FLSAM(stk,tun,ctr,run.dir = "run")
```
Note, if one is using the *flr FLSAM* repository, the above will give the following error (hence it is outcommented):
In the above we get the following error (hence it is outcommented):
```
Error in x@.Data[i, j, k, l, m, n] <- value : replacement has length zero
Called from: `[<-`(`*tmp*`, a, , value = numeric(0))
```
This occurs quite late in the function code within the *SAM2FLR*-function when reading the mortality output matrix. What could be done here is to split the *FLSAM* function into two pieces. Lets source them:
```{r}
source("R/FLSAM_amended.R")
```
Here the FLSAM-function has been split into two sections. One that just does the run (no changes were made in the principle code):
```{r, eval=FALSE}
FLSAM_run(stk,tun,ctr,run.dir="run")
```
Note that the run.dir needs to be specified in the function above.

The second function reads the results back into the *FLR*-environment (using some quick fix):
```{r}
source("R/FLSAM_amended.R")
```
Run the actual assessment:
```{r, eval=FALSE}
FLSAM_run(stk,tun,ctr,run.dir="run")
```

Now we read the data back into the *FLR*-environment (using some quick fix):
```{r, warning=FALSE}
sam <- SAM2FLR_fix(ctrl=ctr,run.dir="run")
# note, here we loose the 2014 data
ass <- stk + sam
```

Lets save the stuff as R-binary:
```{r, eval=FALSE}
save(stk,ctr,sam,ass,file="R/iCod_sam_assessment.RData")
```

## Diagnostics

### Residuals
```{r, warning=FALSE}
res <- residuals(sam)
ggplot(res,aes(year,std.res,fill=std.res>0)) + 
  geom_bar(stat="identity") +
  facet_grid(age ~ fleet) +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Set1") +
  labs(x="",y="")
```

### Observation variance
```{r}
obv <- obs.var(sam)
ggplot(obv,aes(paste(fleet,age))) + 
  geom_pointrange(aes(y=value,ymin=0,ymax=value)) +
  labs(x="",y="") +
  coord_flip()
```

```{r}
otolith(sam,year=2014,plot=T,n=2000)
```

```{r}
cor.plot(sam)
```

```{r}
plot.dat <- catchabilities(sam)
library(lattice)
print(xyplot(value+ubnd+lbnd ~ age | fleet,
data=plot.dat,subset=fleet %in% c("SMB"),
scale=list(alternating=FALSE,y=list(relation="free")),as.table=TRUE,
type="l",lwd=c(2,1,1),col=c("black","grey","grey"),
main="SMB catchability parameters",
ylab="Catchability",xlab="Age"))
```


```{r, eval=FALSE}
residual.diagnostics(sam)
```

### Retro analysis

This does not work:
```{r, eval=FALSE}
retro.sams <- retro(stk,tun,ctr,retro=2)
```

### MCMC

Following does not run:
```{r, eval=FALSE}
x <- monteCarloStock(stk,sam,19,run.dir="run")
```

### Process error
```{r, warning=FALSE}
require(fishvise)
rbya <- FLStock2rbya(ass)
pro <- process_error_sam(rbya)
ggplot(pro,aes(year,dm+m)) + geom_line() + facet_wrap(~ age) +
  geom_hline(yintercept=0.2,col="red") +
  labs(x="",y="")
```

## Results

```{r}
plot(sam)
```

### Comparison with the TRUTH :-)

```{r}
x <- melt(stock.n(sam),value.name = "n")[,c("year","age","n")]
x <- join(x,melt(harvest(sam),value.name = "f")[,c("year","age","f")],by=c("year","age"))
x$run <- "sam"
x$n <- x$n/1e3
truth <- read_adcam("~/ass/2014/01/05ass/Adcam","smb",rName = "truth",ggFactor = FALSE)$rbya
cn <- c("year","age","n","f","run")
x <- rbind(x[,cn],truth[truth$year %in% 1985:2014,cn])
ggplot(x,aes(year,n/1e3,col=run)) + geom_line() +
  facet_wrap(~ age,scale="free_y") +
  theme(legend.position=c(0.8,0.1)) +
  labs(x="",y="",title="Stock in numbers")
ggplot(x,aes(year,f,col=run)) + geom_line() +
  facet_wrap(~ age,scale="free_y") +
  theme(legend.position=c(0.8,0.1)) +
  labs(x="",y="",title="Fishing mortality")
```

```{r, results='asis'}
sessionInfo()
```

## Tables
```{r, results='asis', eval=FALSE}
sam@control@sam.binary <- character()
FLSAM.out(stk,tun,sam)
```